---
title: "calrecycle commercial and residential"
author: "E.M.Thomas"
date: "2024-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(tidyverse)
library(here)
library(janitor)
library(stringr)
library(writexl)
```
# 2014 Waste Characterization Study
### Aggregated by county, material type, and business group / residential 

# Read in busines group data
```{r}
setwd(here("raw data/CalRecycle/Generator-based WCS/Business Group"))

# List all Excel files in your directory
excel_files <- list.files(pattern = "\\.xlsx$", full.names = TRUE)

# Create an empty list to store dataframes
dfs <- list()

# Loop through each Excel file, read it, and append to dfs list
for (file in excel_files) {
  df <- read_excel(file, sheet = 2)
  dfs[[file]] <- df
}

# Combine all dataframes into one
combined_bg_df <- bind_rows(dfs, .id = "file_name")
```

## check for NAs
```{r}
non_na_counties <- combined_bg_df %>%
  select(file_name, `Material Category`) %>% 
  na.omit() 

unique_non_na <- unique(non_na_counties$file_name)

na_counties <- combined_bg_df %>% 
  filter(!file_name %in% unique_non_na) %>% 
  select(file_name)

unique_na <- data.frame(unique(na_counties$file_name)) ### no NA values good
```

## write plastic data to csv
```{r}
### Save the plastic data 
bg_plastic_waste_streams_df <- combined_bg_df %>% 
  clean_names() %>% 
  filter(material_category == "Plastic") %>% 
  select(material_type, jurisdiction_s, material_tons_disposed, material_tons_in_curbside_recycle, material_tons_in_curbside_organics, material_tons_in_other_diversion, material_tons_generated_sum_of_all_streams) %>% 
  rename(county = jurisdiction_s)

#write.csv(bg_plastic_waste_streams_df, "Business Group Waste Stream by County.csv", row.names=FALSE)

bg_plastic_waste_df <- bg_plastic_waste_streams_df %>% 
  select(material_type, county, material_tons_disposed) #changing this to disposed only to match same composition as residential 5/21

#write.csv(bg_plastic_waste_df, "Business Group Waste by County and Material Type.csv", row.names=FALSE)

bg_plastic_sum <- bg_plastic_waste_df %>%
  select(material_type, county, material_tons_disposed) %>% 
  group_by(county) %>%
  summarize(material_tons_disposed = sum(material_tons_disposed))
  
#write.csv(bg_plastic_sum, "Business Group Total Waste Summarized by County.csv", row.names=FALSE)

```


# Read in residential data
```{r}
setwd(here("raw data/CalRecycle/Generator-based WCS/Residential"))

# List all Excel files in your directory
excel_files <- list.files(pattern = "\\.xlsx$", full.names = TRUE)

# Create an empty list to store dataframes
dfs <- list()

# Loop through each Excel file, read it, and append to dfs list
for (file in excel_files) {
  df <- read_excel(file, sheet = 2)
  dfs[[file]] <- df
}

# Combine all dataframes into one
combined_res_df <- bind_rows(dfs, .id = "file_name")
```

## check for NAs
```{r}
non_na_counties <- combined_res_df %>%
  select(file_name, `Material Category`) %>% 
  na.omit() 

unique_non_na <- unique(non_na_counties$file_name)

na_counties <- combined_res_df %>% 
  filter(!file_name %in% unique_non_na) %>% 
  select(file_name)

unique_na <- data.frame(unique(na_counties$file_name)) ### no NA values good
```


## write residential plastic waste to csv
```{r}
### Save the plastic data 
res_plastic_waste_df <- combined_res_df %>% 
  clean_names() %>% 
  filter(material_category == "Plastic") %>% 
  select(material_type, jurisdiction_s, total_residential_tons) %>% 
  rename(county = jurisdiction_s)
  
#write.csv(res_plastic_waste_df, "Residential Total Waste by County.csv", row.names=FALSE)

res_plastic_sum <- res_plastic_waste_df %>%
  group_by(county) %>%
  summarize(total_residential_tons = sum(total_residential_tons))
```

# Merge commericial and residential data
```{r}
# Merge the two data frames based on 'material_type' and 'county'
total_plastic_waste_df <- left_join(bg_plastic_waste_df, res_plastic_waste_df, 
                       by = c("material_type", "county"))

# Calculate the sum of 'material_tons_generated_sum_of_all_streams' and 'total_residential_tons'for every county and material_type observation that is the same between the two data frames
total_plastic_waste_df <- total_plastic_waste_df %>%
  mutate(total_plastic_waste = material_tons_disposed + total_residential_tons) %>%
  select(material_type, county, total_plastic_waste)
```


# Apply Milbrandt conversion formulas

```{r}
### read in resin classification conversion table
resin_conversion_df <- read_excel(here("raw data/Resin Conversion.xlsx")) %>% clean_names()

# Perform a left join between ca_eol_df and resin_conversion_df based on the material_type column
pw_merged_df <- merge(total_plastic_waste_df, resin_conversion_df, by.x = "material_type", by.y = "cal_recycle_classification", all.x = TRUE) %>% 
  mutate(milbrandt_classification = str_trim(milbrandt_classification))
```

## PET
```{r}
### PET 
pet_df <- pw_merged_df %>% 
  group_by(county) %>% 
  summarize(pet = 
              sum(if_else(milbrandt_classification == "PETbc", total_plastic_waste, 
                          if_else(milbrandt_classification == "DP", 0.046 * total_plastic_waste, 
                                  if_else(milbrandt_classification == "RC", 0.08 * total_plastic_waste, 0)))))
```

## HDPE
```{r}
hdpe_df <- pw_merged_df %>% 
  group_by(county) %>% 
  summarize(hdpe = 
              sum(if_else(milbrandt_classification == "HDPEbc", total_plastic_waste, 
                          if_else(milbrandt_classification == "FWB", 0.17 * total_plastic_waste, 
                                  if_else(milbrandt_classification == "DP", 0.114 * total_plastic_waste,
                                          if_else(milbrandt_classification == "RC", 0.12 * total_plastic_waste, 0))))))
```


## PP
```{r}
pp_df <- pw_merged_df %>% 
  group_by(county) %>% 
  summarize(pp = 
              sum(if_else(milbrandt_classification == "MP", 0.65 * total_plastic_waste, 
                          if_else(milbrandt_classification == "FWB", 0.1 * total_plastic_waste, 
                                  if_else(milbrandt_classification == "DP", 0.336 * total_plastic_waste,
                                          if_else(milbrandt_classification == "RC", 0.30 * total_plastic_waste, 0))))))
```

## LDPE/LLDPE
```{r}
ldpe_df <- pw_merged_df %>% 
  group_by(county) %>% 
  summarize(ldpe = 
              sum(if_else(milbrandt_classification == "MP", 0.1 * total_plastic_waste, 
                          if_else(milbrandt_classification == "FWB", 0.69 * total_plastic_waste, 
                                  if_else(milbrandt_classification == "DP", 0.15 * total_plastic_waste,
                                          if_else(milbrandt_classification == "RC", 0.18 * total_plastic_waste, 0))))))
```

## PVC
```{r}
pvc_df <- pw_merged_df %>% 
  group_by(county) %>% 
  summarize(pvc = 
              sum(if_else(milbrandt_classification == "MP", 0.025 * total_plastic_waste, 
                          if_else(milbrandt_classification == "FWB", 0.013 * total_plastic_waste, 
                                  if_else(milbrandt_classification == "DP", 0.018 * total_plastic_waste,
                                          if_else(milbrandt_classification == "RC", 0.036 * total_plastic_waste, 0))))))
```

## Other resins
```{r}
other_resins_df <- pw_merged_df %>% 
  group_by(county) %>% 
  summarize(other_resins = 
              sum(if_else(milbrandt_classification == "DP", 0.279 * total_plastic_waste,
                          if_else(milbrandt_classification == "RC", 0.194 * total_plastic_waste, 0))))
```


## PS
```{r}
ps_df <- pw_merged_df %>% 
  group_by(county) %>% 
  summarize(ps = 
              sum(if_else(milbrandt_classification == "MP", 0.225 * total_plastic_waste, 
                          if_else(milbrandt_classification == "FWB", 0.027 * total_plastic_waste, 
                                  if_else(milbrandt_classification == "DP", 0.057 * total_plastic_waste,
                                          if_else(milbrandt_classification == "RC", 0.09 * total_plastic_waste, 0))))))
```

## Combine resin dfs
```{r}
final_resin_df <- pet_df %>%
  left_join(pp_df, by = "county") %>%
  left_join(pvc_df, by = "county") %>%
  left_join(hdpe_df, by = "county") %>%
  left_join(ldpe_df, by = "county") %>%
  left_join(ps_df, by = "county") %>%
  left_join(other_resins_df, by = "county") %>% 
  mutate(total_plastic_waste = pet + pp + pvc + hdpe + ldpe + ps + other_resins)

final_resin_df$county <- gsub("\\(Countywide\\)$", "", final_resin_df$county)
final_resin_df$county <- str_trim(final_resin_df$county)

#convert short tons to metric tons
final_resin_df <- final_resin_df %>% 
  mutate(across(where(is.numeric), ~ ./1.102))

#write_csv(final_resin_df, "Commercial and Residential Resin Data.csv")
```

## Check the sums 
```{r}
plastic_waste_sums_df <- bg_plastic_sum %>% 
  left_join(res_plastic_sum, by='county') %>% 
  mutate(og_total_plastic_waste = material_tons_disposed + total_residential_tons) %>%
  select(county, og_total_plastic_waste) %>% 
  mutate(across(where(is.numeric), ~ ./1.102))

plastic_waste_sums_df$county <- gsub("\\(Countywide\\)$", "", plastic_waste_sums_df$county)
plastic_waste_sums_df$county <- str_trim(plastic_waste_sums_df$county)

check_plastic_waste_df <- plastic_waste_sums_df %>% 
  left_join(final_resin_df, by = 'county') %>% 
  select(county, og_total_plastic_waste, total_plastic_waste)
```

## Compare w/ interpolated milbrandt
```{r}
milbrandt_interpolated <- read_excel(here("raw data/milbradt_interpolated.xlsx"),2) %>% select(county, resin, mt_interpolated) %>% 
  pivot_wider(names_from = resin, values_from = mt_interpolated) %>% 
  clean_names() %>% 
  select(county, pet_mt, pp_mt, pvc_mt, hdpe_mt, ldpe_mt, ps_mt, other_mt, total_mt)

# Sort dataframes by county
milbrandt_interpolated <- milbrandt_interpolated[order(milbrandt_interpolated$county), ]
final_resin_df <- final_resin_df[order(final_resin_df$county), ]

# Now you can proceed with the subtraction
abs_diff <- milbrandt_interpolated[, -1] - final_resin_df[, -1]

# Extract the "county" column from one of the sorted dataframes (they should be the same)
county_column <- milbrandt_interpolated$county

# Combine the "county" column with abs_diff
abs_diff_with_county <- cbind(county = county_column, abs_diff)

# Calculate the percent difference
percent_diff <- (abs_diff / final_resin_df[, -1]) * 100

# Combine the "county" column with percent_diff
percent_diff_with_county <- cbind(county = county_column, percent_diff)  

```

```{r}
# Write dataframes to the same Excel file
write_xlsx(list("Milbrandt Interpolated Waste Data (metric tons)" = milbrandt_interpolated, "CalRecycle Converted Resin Data (metric tons)" = final_resin_df, "Absolute Difference" = abs_diff_with_county,"Percent Difference" = percent_diff_with_county),
           path = here("processed data/CalRecycle-Milbrandt Comparison_v2.xlsx"),
           col_names = TRUE)
```



